Efficiently stable presentations from error-correcting codes

### 2023-10-02 efficient_stability.tex
minor

Equipped with the new definition we give examples. We first show a general method for constructing candidate stable presentations of $\Z_2^k$ from linear error-correcting codes that have an additional property known as \emph{local testability}. While we cannot show that the method always leads to stable presentations, we leverage recent results in the study of nonlocal games in quantum information theory (Ji et al., Discrete Analysis 2021) to show that a specific instantiation of our construction, based on the Reed-Muller family of codes, leads to a stable presentation of $\Z_2^k$ of size $\poly\log(k)$ only. As an application we combine this result with recent work of de la Salle (arXiv:2204.07084) to re-derive a key building block in the recent disproof of Connes' Embedding Problem via complexity theory (Ji et al., arXiv:2001.04383), the Pauli Braiding Test of Natarajan and Vidick (IEEE FOCS'18). 
\end{abstract}

@@

To summarize, approximate stability results for $G(h)$ enable one to obtain structural results about near-optimal strategies in $\game_h$ --- such a result is known as a \emph{rigidity} result in quantum information. A rigidity result about a game called the Pauli braiding test is at the heart of the proof of $\sf MIP^* = \sf RE$. The analysis of this test requires an efficient stability result for the Weyl-Heisenberg group. Informally, the reason that the stability result needs to be for an efficient presentation of the Weyl-Heisenberg group, as opposed to e.g.\ the multiplication table presentation, is because to obtain the final result it is necessary that the ``complexity'' of the test, or game, is smaller than the ``complexity'' of the object, or group being tested. Of course here we are loose about what we mean by ``complexity,'' and refer to~\cite{ji2021mip,vidickmip} for high-level explanations. The Pauli braiding test is described and analyzed in Section~\ref{sec:pbt}. We end by mentioning an open question: if one was able to obtain $|S_k|,|R_k|=\poly(k)$ while also having $\delta(\eps)=\poly(\eps)$ in the informal theorem stated above, then this result would likely, through the connection we just described, have consequences for the efficient verification of quantum computations in the framework of interactive proof systems---see e.g.~\cite{coladangelo2019verifier,natarajan2023bounding} for a sample of known results in this direction.

@@

\begin{corollary}[Pauli braiding test]
Let $\code_\bRM$ be the $[q^{m+1},t(d+1)^m,D']_2$ Reed-Muller from Section~\ref{sec:eff-z2k}, and $M=(h_\bRM,\nu_\bRM)$ the $(d+2)$-local tester for $\code_\bRM$ described in Figure~\ref{fig:bRM-tester}. Then the associated Pauli braiding test (Figure~\ref{fig:pbt}) is a $(k,\delta')$-qubit test with error function $\delta' = \poly(m,d,t)\cdot\poly(\eps,q^{-1})$.
\end{corollary}

>>>>

Equipped with the new definition we give examples. We first show a general method for constructing candidate stable presentations of $\Z_2^k$ from linear error-correcting codes that have an additional property known as \emph{local testability}. While we cannot show that the method always leads to stable presentations, we leverage recent results in the study of nonlocal games in quantum information theory (Ji et al., Discrete Analysis 2021) to show that a specific instantiation of our construction, based on the Reed-Muller family of codes, leads to a stable presentation of $\Z_2^k$ of size $\poly\log(k)$ only. As an application we combine this result with recent work of de la Salle (arXiv:2204.07084) to re-derive a key building block in the recent disproof of Connes' Embedding Problem via complexity theory (Ji et al., arXiv:2001.04383), the quantum low-degree test of Natarajan and Vidick (IEEE FOCS'18). 
\end{abstract}

@@

To summarize, approximate stability results for $G(h)$ enable one to obtain structural results about near-optimal strategies in $\game_h$ --- such a result is known as a \emph{rigidity} result in quantum information. A rigidity result about a game called the quantum low-degree test~\cite{natarajan2018low} is at the heart of the proof of $\sf MIP^* = \sf RE$. The analysis of this test requires an efficient stability result for the Weyl-Heisenberg group. Informally, the reason that the stability result needs to be for an efficient presentation of the Weyl-Heisenberg group, as opposed to e.g.\ the multiplication table presentation, is because to obtain the final result it is necessary that the ``complexity'' of the test, or game, is smaller than the ``complexity'' of the object, or group being tested. Of course here we are loose about what we mean by ``complexity,'' and refer to~\cite{ji2021mip,vidickmip} for high-level explanations. The quantum low-degree test is described and analyzed in Section~\ref{sec:pbt}. We end by mentioning an open question: if one was able to obtain $|S_k|,|R_k|=\poly(k)$ while also having $\delta(\eps)=\poly(\eps)$ in the informal theorem stated above, then this result would likely, through the connection we just described, have consequences for the efficient verification of quantum computations in the framework of interactive proof systems---see e.g.~\cite{coladangelo2019verifier,natarajan2023bounding} for a sample of known results in this direction.

@@

\begin{corollary}[Quantum low-degree test]
Let $\code_\bRM$ be the $[q^{m+1},t(d+1)^m,D']_2$ Reed-Muller from Section~\ref{sec:eff-z2k}, and $M=(h_\bRM,\nu_\bRM)$ the $(d+2)$-local tester for $\code_\bRM$ described in Figure~\ref{fig:bRM-tester}. Then the associated braiding test (Figure~\ref{fig:pbt}) is a $(k,\delta')$-qubit test with error function $\delta' = \poly(m,d,t)\cdot\poly(\eps,q^{-1})$.
\end{corollary}
